### 2017-12-14 author.tex
collect common affiliations to avoid spelling errors

\newcommand{\afprinceton}{Department of Astrophysical Sciences, Princeton University, Princeton, NJ 08544, USA}
\newcommand{\afstsci}{Space Telescope Science Institute, 3700 San Martin Dr., Baltimore, MD 21218, USA}
\newcommand{\afsaao}{South African Astronomical Observatory, PO Box 9, Observatory 7935, Cape Town, South Africa}
\newcommand{\afminnstate}{Department of Physics and Astronomy, Minnesota State University Moorhead, 1104 7th Ave S, Moorhead, MN 56563}
\newcommand{\afgoddard}{NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
\newcommand{\afcfa}{Harvard-Smithsonian Center for Astrophysics, 60 Garden St., Cambridge, MA, 02138, USA}
\newcommand{\afwesternontario}{Department of Physics \& Astronomy, University of Western Ontario, 1151 Richmond St, London ON N5X4H1 Canada}
\newcommand{\afnasaames}{NASA Ames Research Center, Moffett Field, CA 94043, USA}
\newcommand{\afjhu}{Department of Physics and Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}
\newcommand{\afmpik}{Max-Planck-Institut f\"ur Kernphysik, PO Box 103980, 69029 Heidelberg, Germany}
\newcommand{\afioa}{Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
\newcommand{\afpennstate}{Dept of Astronomy and Astrophysics, Pennsylvania State University, University Park, PA 16802}
\newcommand{\afgemini}{Gemini Observatory, 670 N. Aohoku Pl, Hilo, HI 96720, USA}
\newcommand{\aflco}{Las Cumbres Observatory, 6740 Cortona Drive, Suite 102, Goleta, CA 93117-5575, USA}
\newcommand{\afucsb}{Department of Physics, University of California, Santa Barbara, CA 93106-9530, USA}
\newcommand{\afesomunich}{European Southern Observatory, Karl-Schwarzschild-Stra{\ss}e 2, 85748 Garching bei M\"{u}nchen, Germany}
\newcommand{\afuw}{Department of Astronomy, University of Washington, Seattle, WA 98155}
\newcommand{\afberkeleyastro}{Department of Astronomy, UC Berkeley, 501 Campbell Hall \#3411, Berkeley, CA 94720, USA}

\collaboration{Primary paper contributors:}

\author[0000-0003-0872-7098]{A. M. Price-Whelan}
\affiliation{\afprinceton}

\author[0000-0002-3713-6337]{B. M. Sip\H{o}cz}
\affiliation{}

\author[0000-0003-4243-2840]{H. M. G\"{u}nther}
\affiliation{Kavli Institute for Astrophysics and Space Research, Massachusetts Institute of Technology, 70 Vassar St., Cambridge, MA 02139, USA}

\author[0000-0003-0079-4114]{P. L. Lim}
\affiliation{\afstsci}

\author[0000-0002-8969-5229]{S. M. Crawford}
\affiliation{\afsaao}

\author{S. Conseil}
\affiliation{Univ Lyon, Univ Lyon1, Ens de Lyon, CNRS, Centre de Recherche Astrophysique de Lyon UMR5574, F-69230, Saint-Genis-Laval, France}

\author[0000-0003-4401-0430]{D. L. Shupe}
\affiliation{Caltech/IPAC, 1200 E. California Blvd, Pasadena, CA 91125}

\author[0000-0001-7988-8919]{M. W. Craig}
\affiliation{\afminnstate}

\author[0000-0002-5686-9632]{N. Dencheva}
\affiliation{\afstsci}

\author[0000-0001-6431-9633]{A. Ginsburg}
\affiliation{National Radio Astronomy Observatory, 1003 Lopezville Rd, Socorro, NM 87801}

\author[0000-0002-9623-3401]{J. T. VanderPlas}
\affiliation{eScience Institute, University of Washington, 3910 15th Ave NE, Seattle, WA 98195, USA}

\author[0000-0002-7908-9284]{L. D. Bradley}
\affiliation{\afstsci}

\author[0000-0003-0784-6909]{D. P\'{e}rez-Su\'{a}rez}
\affiliation{University College London/Research IT Services, Gower St, Bloomsbury, London WC1E 6BT, United Kingdom}

\author[0000-0002-0455-9384]{M. de Val-Borro}
\affiliation{Astrochemistry Laboratory, \afgoddard}

\collaboration{Astropy coordination committee:}

\author{T. L. Aldcroft}
\affiliation{\afcfa}

\author[0000-0002-1821-0650]{K. L. Cruz}
\affiliation{Department of Physics and Astronomy, Hunter College, City University of New York, 695 Park Avenue, New York, NY 10065}
\affiliation{Physics, Graduate Center of the City University of New York, New York, NY, USA}
\affiliation{Department of Astrophysics, American Museum of Natural History, New York, NY, USA}

\author[0000-0002-8642-1329]{T. P. Robitaille}
\affiliation{Aperio Software Ltd., Headingley Enterprise and Arts Centre, Bennett Road, Leeds, LS6 3HN, United Kingdom}

\author[0000-0002-9599-310X]{E. J. Tollerud}
\affiliation{\afstsci}

\collaboration{Astropy contributors:}

\author{C. Ardelean}
\affiliation{\afwesternontario}

\author{T. Babej}
\affiliation{Department of Theoretical Physics \& Astrophysics, Masaryk University, Kotlarska 2, 61137 Brno, Czech Republic}

\author[0000-0002-4576-9337]{M. Bachetti}
\affiliation{INAF-Osservatorio Astronomico di Cagliari, via della Scienza 5, I-09047, Selargius, Italy}

\author{A. V. Bakanov}
\affiliation{}

\author[0000-0001-7821-7195]{S. P. Bamford}
\affiliation{School of Physics \& Astronomy, University of Nottingham, University Park, Nottingham NG7 2RD, UK}

\author[0000-0002-3306-3484]{G. Barentsen}
\affiliation{\afnasaames}

\author[0000-0003-2767-0090]{P. Barmby}
\affiliation{\afwesternontario}

\author[0000-0002-9374-2729]{A. Baumbach}
\affiliation{Heidelberg University, Kirchhoff Institut for Physics, Im Neuenheimer Feld 227, 69116 Heidelberg, Germany}

\author{K. L. Berry}
\affiliation{}

\author{F. Biscani}
\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\author[0000-0003-0946-6176]{M. Boquien}
\affiliation{Unidad de Astronomía, Fac. Cs. Básicas, Universidad de Antofagasta, Avda. U. de Antofagasta 02800, Antofagasta, Chile}

\author{K. A. Bostroem}
\affiliation{Department of Physics, UC Davis, 1 Shields Ave, Davis, CA, 95616, USA}

\author{L. G. Bouma}
\affiliation{\afprinceton}

\author[0000-0003-2680-005X]{G. B. Brammer}
\affiliation{\afstsci}

\author{E. M. Bray}
\affiliation{}

\author{H. Breytenbach}
\affiliation{\afsaao}

\author[0000-0001-8001-0089]{H. Buddelmeijer}
\affiliation{Leiden Observatory, Leiden University, P.O. Box 9513, 2300 RA, Leiden, The Netherlands}

\author[0000-0003-4428-7835]{D. J. Burke}
\affiliation{\afcfa}

\author[0000-0002-7738-5389]{G. Calderone}
\affiliation{Istituto Nazionale di Astrofisica, via Tiepolo 11 Trieste, Italy}

\author{M. Cara}
\affiliation{\afstsci}

\author{J. V. M. Cardoso}
\affiliation{Universidade Federal de Campina Grande, Campina Grande, PB 58429-900, Brazil}
\affiliation{\afnasaames}
\affiliation{Bay Area Environmental Research Institute, Petaluma, CA 94952, USA}

\author{S. Cheedella}
\affiliation{Department of Physics, Virginia Tech, Blacksburg, VA 24061, USA}

\author[0000-0002-5317-7518]{Y. Copin}
\affiliation{CNRS/IN2P3, Institut de Physique Nucleaire de Lyon, Universite de Lyon, Universite de Lyon 1, Villeurbanne, F-69622, Lyon, France}

\author[0000-0003-1204-3035]{D. Crichton}
\affiliation{\afjhu}

\author{D. D'Avella}
\affiliation{\afstsci}

\author[0000-0002-4198-4005]{C. Deil}
\affiliation{\afmpik}

\author[0000-0003-0526-3873]{\'{E}. Depagne}
\affiliation{\afsaao}

\author[0000-0002-8134-9591]{J. P. Dietrich}
\affiliation{Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 181679 Munich, Germany}
\affiliation{Excellence Cluster Universe, Boltzmannstr. 2, 85748 Garching b. M\"unchen, Germany}

\author{A. Donath}
\affiliation{\afmpik}

\author{T. Erben}
\affiliation{Argelander-Institut f\"ur Astronomie, Auf dem H\"ugel 71, 53121 Bonn, Germany}

\author[0000-0002-8919-079X]{L. A. Ferreira}
\affiliation{Instituto de Matemática Estatística e Física – IMEF, Universidade Federal do Rio Grande – FURG, Rio Grande, RS 96203-900, Brazil}

\author{T. Finethy}
\affiliation{}

\author[0000-0003-4291-1091]{R. T. Fox}
\affiliation{}

\author[0000-0002-9853-5673]{L. H. Garrison}
\affiliation{\afcfa}

\author{S. L. J. Gibbons}
\affiliation{\afioa}

\author{D. A. Goldstein}
\affiliation{\afberkeleyastro}
\affiliation{Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA}

\author[0000-0002-0300-3333]{R. Gommers}
\affiliation{Scion, Private Bag 3020, Rotorua, New Zealand}

\author[0000-0003-4970-2874]{J. P. Greco}
\affiliation{\afprinceton}

\author{P. Greenfield}
\affiliation{\afstsci}

\author[0000-0002-6508-2938]{A. M. Groener}
\affiliation{Drexel University, Physics Department, Philadelphia, PA 19104, USA}

\author{F. Grollier}
\affiliation{}

\author[0000-0003-2031-7737]{A. Hagen}
\affiliation{Vizual.ai, 3600 O'Donnell St, Suite 250, Baltimore, MD 21224}
\affiliation{\afpennstate}

\author{P. Hirst}
\affiliation{\afgemini}

\author[0000-0002-8546-9128]{D. Homeier}
\affiliation{Zentrum f{\"u}r Astronomie der Universit{\"a}t Heidelberg, Landessternwarte, K{\"o}nigstuhl 12, D-69117 Heidelberg, Germany}

\author[0000-0002-4600-7852]{A. J. Horton}
\affiliation{Australian Astronomical Observatory, 105 Delhi Road, North Ryde NSW 2113, Australia}

\author[0000-0002-0832-2974]{G. Hosseinzadeh}
\affiliation{\aflco}
\affiliation{\afucsb}

\author{L. Hu}
\affiliation{Imperial College London,  Kensington, London SW7 2AZ, United Kingdom}

\author[0000-0003-4989-0289]{J. S. Hunkeler}
\affiliation{\afstsci}

\author[0000-0001-5250-2633]{\v{Z}. Ivezi\'{c}}
\affiliation{\afuw}

\author{A. Jain}
\affiliation{BITS PILANI/Computer Science, Pilani Campus, Rajasthan, India}

\author[0000-0001-5982-167X]{T. Jenness}
\affiliation{Large Synoptic Survey Telescope, 950 N Cherry Ave, Tucson, AZ, 85719}

\author{G. Kanarek}
\affiliation{\afstsci}

\author[0000-0002-7612-0469]{S. Kendrew}
\affiliation{European Space Agency, \afstsci}

\author[0000-0002-8211-1892]{N. S. Kern}
\affiliation{\afberkeleyastro}

\author[0000-0002-0479-7235]{W. E. Kerzendorf}
\affiliation{\afesomunich}

\author{A. Khvalko}
\affiliation{}

\author{J. King}
\affiliation{\afmpik}

\author[0000-0002-8828-5463]{D. Kirkby}
\affiliation{Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA}

\author{A. M. Kulkarni}
\affiliation{College of Engineering Pune/Department of Computer Engineering and IT, Shivajinagar, Pune 411005, India}

\author{A. Kumar}
\affiliation{Delhi Technological University}

\author[0000-0003-2193-5369]{A. Lee}
\affiliation{Department of Physics, University of Berkeley, Califonia, CA94709, USA}

\author[0000-0001-5820-475X]{D. Lenz}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Pasadena, CA 91109, USA}

\author[0000-0001-7221-855X]{S. P. Littlefair}
\affiliation{Department of Physics \& Astronomy, University of Sheffield, Sheffield, S3 7RH, UK}

\author[0000-0003-3270-6844]{Z. Ma}
\affiliation{Department of Physics and Astronomy, University of Missouri, Columbia, Missouri, 65211, USA}

\author[0000-0002-1395-8694]{D. M. Macleod}
\affiliation{Cardiff University, Cardiff CF24 3AA, UK}

\author[0000-0002-6324-5713]{M. Mastropietro}
\affiliation{Department Physics and Astronomy, Ghent University, Krijgslaan 281, S9, B-9000 Gent, Belgium}

\author[0000-0001-5807-7893]{C. McCully}
\affiliation{\aflco}
\affiliation{\afucsb}

\author{S. Montagnac}
\affiliation{Puy-Sainte-R\'eparade Observatory}

\author[0000-0003-2528-3409]{B. M. Morris}
\affiliation{\afuw}

\author[0000-0003-4217-4642]{S. J. Mumford}
\affiliation{SP$^{2}$RC, School of Mathematics and Statistics, The University of Sheffield, U.K.}

\author[0000-0002-1631-4114]{D. Muna}
\affiliation{Center for Cosmology and Astroparticle Physics, The Ohio State University, 191 West Woodruff Avenue, Columbus, OH 43210}

\author[0000-0001-6628-8033]{N. A. Murphy}
\affiliation{\afcfa}

\author{S. Nelson}
\affiliation{\afminnstate}

\author[0000-0002-1966-3627]{G. H. Nguyen}
\affiliation{VNU-HCMC, University of Natural Sciences/Faculty of IT, 227 Nguyen Van Cu St., Ward 4, District 5, Ho Chi Minh City, Vietnam}

\author[0000-0001-8720-5612]{J. P. Ninan}
\affiliation{\afpennstate}

\author{M. N{\"o}the}
\affiliation{TU Dortmund, Experimental Physics 5, Otto-Hahn-Str. 4, 44221 Dortmund, Germany}

\author{S. Ogaz}
\affiliation{\afstsci}

\author[0000-0001-7790-5308]{S. Oh}
\affiliation{\afprinceton}

\author{J. K. Parejko}
\affiliation{\afuw}

\author{N. Parley}
\affiliation{University of Reading, Whiteknights Campus, Reading RG6 6BX, UK}

\author[0000-0002-9351-6051]{S. Pascual}
\affiliation{Departamento de Astrofisica, Universidad Complutense de Madrid, Madrid, Spain}

\author{R. Patil}
\affiliation{}

\author{A. A. Patil}
\affiliation{Pune Institute of Computer Technology, Pune 411043, India}

\author[0000-0002-9912-5705]{A. L. Plunkett}
\affiliation{European Southern Observatory, Av. Alonso de C\'{o}rdova 3107, Vitacura, Santiago, Chile}

\author{T. Rastogi}
\affiliation{}

\author{V. Reddy Janga}
\affiliation{Indian Institute of Technology, Mechanical Engineering, Kharagpur, India}

\author[0000-0003-1149-6294]{J. Sabater}
\affiliation{Institute for Astronomy (IfA), University of Edinburgh, Royal Observatory, Blackford Hill, EH9 3HJ Edinburgh, U.K.}

\author{P. Sakurikar}
\affiliation{IIIT-Hyderabad, India}

\author{M. Seifert}
\affiliation{}

\author{L. E. Sherbert}
\affiliation{\afstsci}

\author[0000-0003-0477-6220]{H. Sherwood-Taylor}
\affiliation{}

\author{A. Y. Shih}
\affiliation{\afgoddard}

\author{M. T. Silbiger}
\affiliation{}

\author[0000-0003-2462-7273]{S. Singanamalla}
\affiliation{Microsoft Research}

\author[0000-0001-9898-5597]{L. P. Singer}
\affiliation{Astroparticle Physics Laboratory, \afgoddard}

\author[0000-0003-1585-225X]{P. H. Sladen}
\affiliation{ARI/ZAH, Heidelberg}

\author{K. A. Sooley}
\affiliation{}

\author{S. Sornarajah}
\affiliation{}

\author[0000-0001-7751-1843]{O. Streicher}
\affiliation{Leibniz Institute for Astrophysics Potsdam (AIP), An der Sternwarte 16, 14482 Potsdam, Germany}

\author[0000-0003-1774-3436]{P. Teuben}
\affiliation{Astronomy Department, University of Maryland, College Park, MD. 20742, USA}

\author{S. W. Thomas}
\affiliation{\afioa}

\author{J. E. H. Turner}
\affiliation{\afgemini}

\author{V\'{i}ctor Terr\'{o}n}
\affiliation{Institute of Astrophysics of Andalusia (IAA-CSIC), Spain}

\author[0000-0002-5830-8505]{M. H. van Kerkwijk}
\affiliation{Department of Astronomy \& Astrophysics, University of Toronto, 50 Saint George Street, Toronto, ON M5R 3B3, Canada}

\author[0000-0002-6219-5558]{A. de la Vega}
\affiliation{\afjhu}

\author[0000-0002-1343-134X]{L. L. Watkins}
\affiliation{\afstsci}

\author{B. A. Weaver}
\affiliation{National Optical Astronomy Observatory, 950 N. Cherry Ave., Tucson, AZ 85719}

\author[0000-0003-4824-2087]{J. B. Whitmore}
\affiliation{Centre for Astrophysics and Supercomputing, Swinburne University of Technology, Hawthorn, VIC 3122, Australia}

\author[0000-0002-2958-4738]{J. Woillez}
\affiliation{\afesomunich}

\author{J. Xavier Prochaska}
\affiliation{Astronomy \& Astrophysics, UC Santa Cruz, 1156 High St., Santa Cruz, CA 95064 USA}

\author[0000-0003-2638-7648]{V. Zabalza}
\affiliation{}