### 2018-03-02 response.txt

We thank the referee for the close reading of the paper and the insightful
comments.  We provide a response to specific comments from the referee
and ths statistical editor below.

> There are some stylistic issues with the text. I have left comments inline
>  where appropriate.

> More generally, given the technical nature of the work being described here,
> I frequently found myself wanting a small code snippet that demonstrated the
> new/improved functionality of the packages being described. Perhaps the
> authors could find a way to include short code examples where appropriate,
> either inline in the paper body or in the appendix.

> Specific issues that need to be corrected: There are some errors in the paper
> (inline comments left where appropriate), for example, Figure 6 has a fairly
> major issue with either the wrong caption or the wrong figure but I'm sure
> this is easily addressed. I have left comments throughout the modified
> 'main.tex' file in the form `\inlinecomment{REVIEWER}{Comment}`

> More generally I have a few comments/concerns:

> Given the title of this paper, I expected to see more discussion by the
> authors about the inclusivity of their community. An uber-skeptical reader
> could make the assessment that the Astropy community has a code of conduct
> therefore believes it's inclusive. I suspect that the community is doing
> much more than this though... So, has any attempt been made by the authors
> to quantify the inclusivity of the Astropy community? There's a growing
> literature in the quantitative social sciences studying open source
> communities that the authors might want to look into: e.g. the work of
> Gousios http://www.gousios.gr/publications.html and Vasilescu
> https://cmustrudel.github.io/publications/ come to mind. Some discussion
> or attempt to quantify the inclusivity is recommended.

We thank the referee for bringing up this important point. After some discussion
within the Astropy community, it became clear that a section regarding inclusion
is beyond the scope of the current paper. Quantitative analysis of the Astropy
community demographics is warranted. However, this is a serious undertaking
beyond the capabilities of our community, and it may require IRB approval. Based
on the referee's comment, we have established initial contact with the NumFOCUS
Diversity & Inclusion in Scientific Computing (DISC) program. Results will not
be available in time for this paper. We have opted to remove "inclusive" from
the paper title to make room for future study and dialogue.

> The phase 'open development' is used a few times but isn't really defined or
> connected to the broader open source ecosystem.  What other communities are
> working in this way? How novel is 'open development' by their assessment? For
> example, other communities have adopted policies that encourage community
> contribution - most notable is the Node.js Foundation's contribution policies
> designed to promote 'Healthy Open Source'
> https://medium.com/the-node-js-collection/healthy-open-source-967fa8be7951 .
> How does the Astropy community efforts compare to this?

> A number of times I found myself wondering how the Astropy projects decides
> whether to build their own package or leverage something available in the
> scientific Python ecosystem? On the face of it it seems like every decision
> made by the Astropy project to build a new core package only increases the
> long term cost/overhead of maintaining this functionality. For example,
> Section 3.5.2 describes `astropy.table` and notes that there's similarities
> to the pandas package and numpy arrays. The functionality described in
> `astropy.nddata` also sounds somewhat generic. Some discussion as to why
> existing community tools (such as pandas) weren't used would be very welcome.
> In addition, if the answer to this question is that these other tools don't
> support all of the functionality the Astropy community needs (e.g. ability to
> handle Quantities with units in tables) are any efforts being made by the
> Astropy community to contribute these changes to packages such as Pandas?

This is an important point, but turns out to be very difficult to generalize how these decisions are made (ironically!). From a *dependency* point of view, Astropy deeply intertwines with the existing Python scientific ecosystem. Where possible we always try to use other tools instead of building our own. Some major astropy functionality is essentially an astro-relevant "wrapper" around existing code. For example, modeling makes it easier to use fitting functions from scipy with models that are primarily of interest in astronomy. Other bits of functionality are really astro-specific or the work to make it more "general" is difficult.  The coordinates and nddata subpackages are good examples of this latter case - despite the general-sounding names, they are focused on the needs of the astronomy community.  The `nddata` case the referee mentions is particularly notable because the actual underlying machinery is built completely on `numpy`, but essentially provides a standard "metadata" language for N-dimensional data *in astronomy*.  That is, it has attributes like `wcs`, `flags`, etc, that have a particular meaning and shared "cultural" understanding in astronomy that is lost if it is made into a more "general" tool.

Other pieces were originally more astro-specific or had no equivalent when they were started and have grown more generalizable as time goes on - e.g. units and table. In general, it's considered "best effort" to generalize these things because the first priority is basic tools for astronomy. Such changes come with non-trivial added cost because that means it now needs a maintainer to do releases, review pull requests, maintain the docs, etc. Using the astropy community helps this, of course, but it is still more work than simply leaving these things in the core. So how much anything is generalized dependes heavily on the time/interest of an individual to do the porting. So for example some of the documentation and testing machinery is now (after v2.0) getting moved from astropy into more generally-useful independent packages.  Meanwhile, units/Quantity would require a much larger effort to disentangle from astropy and maintain independently - but no maintainer with enough time has thus far volunteered.

We have added a new section 2.7 outlining some of the above discussion.

Additionally, as the referee points out, `astropy.table` represents a particularly unusual case for this.  It falls under the "historical" case in that it was started before Pandas.  By the time Pandas became a standard tool in the pydata stack, tables had already developed a fairly large user base in astronomy, so it was not an option to completely eliminate it and say "use Pandas".  It had also developed several features missing from Pandas but important for astronomy users (e.g., support for units, fully-fledged objects as columns, and support for the FITS table concept of "multidimensional columns").  The `astropy.table` investigated this possibility, but it was clear that some of the features would require major incompatible changes *in Pandas* to do this, and it was clear by then that the Pandas user base was too large to accept this.  This led to the current solution of `to_pandas` and `from_pandas`, which allows transformation back-and-forth for the use cases where this is supported. We have added a briefer summary of this discussion to the text in the tables subsection (3.5.2).

________________________

> Note from statistical editor.

> In sec 3.11.3, a sentence can be added warning users about difficulties in
> interpreting the statistical significance of Lomb -Scargle periodogram peaks.
> This is important because unreliable interpretation is common, and warnings
> are not given elsewhere (e.g. NASA Exoplanet Archive Periodogram Service).
> A good reference is Vanderplas 2017 (https://arxiv.org/abs/1703.09824)
> soon to appear in the ApJ Suppl.

A short note directing the reader to Vanderplas (2007) was included.